---
title: "Project Visuals"
description: |
  During the fall semester of 2021, I completed an introductory R course. This post includes graphs and visuals using my final project dataset. The [dataset](https://osf.io/kthnf/) was last updated in October of 2020 and focuses on pregnancies, births, and abortions in the United States for the years 1973-2017. This dataset was compiled and analyzed by Isaac Maddow-Zimet, Kathryn Kost, and Sean Finn. Within my project, I focused on pregancy rates from 2000-2017 for four different age groups: 20-24, 25-29, 30-34, and 35-39.
author:
  - name: Isabella Mahmoud
date: 2022-01-24
output:
  distill::distill_article:
    self_contained: false
---

```{r}
library(tidyverse)
library(gt)
library(janitor)
library(ggplot2)
library(readr)
Birth <- read_csv("~/STA 518/Activities/Final Mess Around/NationalAndStatePregnancy_PublicUse.csv")
```

```{r setup, include=FALSE}
Birth %>%
select(state, year, pregnancyrate2024:pregnancyrate3539) %>%
filter(year >= 2000, state !="US" & state != "DC") %>%
clean_names(case=c("upper_camel")) %>%
rename(
  "Pregnancy Rate 20-24" = Pregnancyrate2024,
  "Pregnancy Rate 25-29" = Pregnancyrate2529,
  "Pregnancy Rate 30-34" = Pregnancyrate3034,
  "Pregnancy Rate 35-49" = Pregnancyrate3539
) %>%
gt() %>%
  cols_width (
    State ~ px(60),
    ends_with("r") ~ px(60),
    everything() ~ px(180)
)
Birth
```

This chunk is filtering the data to what is being focused on in this project.

```{r}
Birth2 <- Birth %>%
  select("state", "year", "pregnancyrate2024":"pregnancyrate3539") %>%
  filter(year >=2000, state !="US" & state != "DC")
Birth2
```

This code creates a bar graph displaying the pregnancy rates for all age groups for the year 2000.

```{r}
Bir <- Birth2 %>%
  filter(year==2000) %>%
  pivot_longer(c("pregnancyrate2024":"pregnancyrate3539"), names_to = "Groups", values_to = "Rates") %>%
  group_by(Groups, year) %>%
  summarise(Rate_Avg = mean(Rates)) %>%
  ggplot() +
  geom_col(aes(x=Groups, y=Rate_Avg, fill=Groups)) +
  scale_x_discrete(labels=c("20-24", "25-29", "30-34", "35-39")) +
  labs(x = "Age groups",
       y= "Average pregnancy rates",
       title = "Average pregnancy rates by age for the year 2000 in the United States",
       caption = "*Rates are per 1,000 women",
       tag = "Fig. 1")  +
  theme(axis.text.x = element_text(angle = 60, size=12, hjust = 1))
Bir
```

The code chunk creates a bar graph for the year 2017 for average pregnancy rates among all age groups.

```{r}
Bir2 <- Birth2 %>%
  filter(year==2017) %>%
  pivot_longer(c("pregnancyrate2024":"pregnancyrate3539"), names_to = "Groups", values_to = "Rates") %>%
  group_by(Groups, year) %>%
  summarise(Rate_Avg = mean(Rates)) %>%
ggplot() +
  geom_col(aes(x=Groups, y=Rate_Avg, fill=Groups)) +
  scale_x_discrete(labels=c("20-24", "25-29", "30-34", "35-39")) +
  labs(x = "Age groups",
       y= " Average pregnancy rates",
       title = "Average pregnancy rates by age for the year 2017 in the United States",
       caption = "*Rates are per 1,000 women",
       tag = "Fig. 2") +
  theme(axis.text.x = element_text(angle = 60, size=12, hjust = 1))
Bir2
```

In order to create a linge graph displaying the birth rates for each year for each age group, I had to use the pivot_longer function.
This allowed me to attach rates to different age groups per year. Once grouping by groups and year, I used "summarise" to get an average rate. Doing this made it so I could graph group averages per year over a 17 year period.

```{r}
Bir3 <- Birth2 %>%
    pivot_longer(c("pregnancyrate2024":"pregnancyrate3539"), names_to = "Groups", values_to = "Rates") %>%
    group_by(Groups, year) %>%
    summarise(Rate_Avg = mean(Rates)) %>%
    ggplot(mapping = aes(x = year, y = Rate_Avg, color = Groups)) + 
    geom_point()+
    geom_line() +
  labs(x = "Year",
       y= "Average pregnancy rates",
       title = "Average pregnancy rates by year in the United States from 2000-2017",
       caption = "*Rates are per 1,000 women",
       tag = "Fig. 3")
Bir3
```

This next part is looking at abortion rates for 20-24 year olds in Illinois over 17 years. Since I'm from Illinois and within this age group, I thought this data would be most relevant/interesting to me. 

```{r}
rib <- Birth %>%
  select("state", "year", "abortionrate2024") %>%
  filter(year >=2000, state == "IL") %>%
  pivot_longer(c("abortionrate2024"), names_to = "Groups", values_to = "Rates")
rib
```

This chunk creates a line graph for the code chunk above. 

```{r}
rib %>%
  ggplot(mapping=aes( x=year, y=Rates)) +
    geom_point() +
    geom_smooth(method=lm, se=F) +
   labs(x = "Year",
       y= "Average abortion rates",
       title = "Average abortion rates for 20-24 year olds per year in Illinois from 2000-2017",
       caption = "*Rates are per 1,000 women",
       tag = "Fig. 4")
```

A model is made based off the simple linear regression created above. Broom::tidy focuses on specific statistics, such as the coefficents for the intercept and predictor, standard error, test statistic, and p-value. 

```{r}
model <- lm(Rates ~ year, data=rib)
summary(model)

#this will make things look nicer
model %>% 
  broom::tidy()
```



